Optimal (Multiway) Spatial Joins

### 2024-05-19 msj.tex
yufei

\extraspacing {\bf Math Conventions.}

@@

From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time with planesweep \cite{bcko08}. However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario. All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ... \times R_k$.

>>>>

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle. \todo{more?}

@@

From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time using $O(n)$ space (see, e.g., \cite{bcko08}). However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario. All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ... \times R_k$.

\begin{table*} 
    \begin{tabular}{c|c|c|c|c} 
        $\bm{k}$ & {\bf method} & {\bf runtime} (big-O) & {\bf space} (big-O) & {\bf remark} \\
        \hline\hline 
        2 & folklore & $n \log n + \out$ & $n$ & comparison-based optimal \\ 
        \hline
        $\ge 3$ & before 2022 & $n^k$ & - & \\ 
        $\ge 3$ & \cite{ty22} & $(n + \out) \cdot \polylog n$ & $n \polylog n$ & \\ 
        $\ge 3$ & \cite{kcko22} & $n \log^{2k} n + \out$ & $n \log n$ & \\ 
        \hline
        $\ge 3$ & ours & $n \log n + \out$ & $n$ & comparison-based optimal
    \end{tabular}

    \vspace{3mm}
    \caption{Comparison of results on the $k$-SJ problem} 
    \label{tab:results-com}
\end{table*}
